Dynamics of Colombo's Top: Tidal Dissipation and Resonance Capture

### 2020-08-18 writeups/Su_disk_disp.tex
done response

In this paper, we consider a wide range of initial planetary obliquities in the
Millholland-Batygin dissipating disk scenario, and examine how the obliquity
evolves toward the ``final'' value as the exterior disk dissipates. We provide
an analytical framework for understanding the final planetary obliquity for
arbitrary initial spin-disk misalignment angles. We also consider various
dissipation timescales, and examine both ``adiabatic'' (slow disk dissipation)
and ``non-adiabatic'' evolution. \textcolor{Corr}{Our chief new results include
that: (i) careful accounting of phase space area across separatrix crossings can
analytically describe the rich dynamics of adiabatic evolution, and (ii) the
concept of ``partial adiabatic resonance advection'' fully captures dynamics in
the non-adiabatic limit.} We calibrate our analytical results with numerical
calculations.

@@

The spin axis $\uv{s}$ of the planet tends to precess around its orbital
(angular momentum) axis $\uv{l}$, driven by the gravitational torque from the
host star acting on the planet's rotational bulge. On the other hand, $\uv{l}$
and the disk axis $\uv{l}_{\rm d}$ precess around each other due to
gravitational interactions. We assume $S \ll L \ll L_{\rm d}$, so $\uv{l}_{\rm
d}$ is nearly constant and $\uv{l}$ experiences negligible backreaction torque
from $\uv{s}$. The equations of motion for $\uv{s}$ and $\uv{l}$ in this limit
are \textcolor{Corr}{\citep{anderson2018teeter}}
\begin{align}
    \rd{\uv{s}}{t} &= \omega_{\rm sl} \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}}
        \equiv \alpha \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}},\label{eq:dsdt}\\
    \rd{\uv{l}}{t} &= \omega_{\rm ld}\p{\uv{l} \cdot \uv{l}_{\rm d}}
            \p{\uv{l} \times \uv{l}_{\rm d}}
        \equiv -g\p{\uv{l} \times \uv{l}_{\rm d}},
            \label{eq:dldt}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &\equiv \frac{3GJ_2 M_pR_p^2 M_\star}{2a_p^3 I_p\Omega_p}
        = \frac{3k_{\rm qp}}{2k_{\rm p}} \frac{M_\star}{m_{\rm p}}
            \p{\frac{R_{\rm p}}{a_{\rm p}}}^3 \Omega_{\rm p},\label{eq:wsl}\\
    \omega_{\rm ld} &\equiv \frac{3M_{\rm d}}{4M_\star}\p{\frac{a_{\rm
            p}}{r_{\rm d}}}^3 n .\label{eq:wld}
\end{align}
In Eq.~\eqref{eq:wsl}, $I_p = k_p M_pR_p^2$ (with $k_p$ a constant) is the
moment of inertia and $J_2=k_{\rm qp}\Omega_p^2 (R_p^3/GM_p)$ (with $k_{qp}$ a
constant) the rotation-induced (dimensionless) quadrupole of the planet [for a
body with uniform density, $k_p=0.4, k_{qp}=0.5$; for giant planets, $k_p\simeq
0.25$ and $k_{qp}\simeq 0.17$ \citep[e.g.][]{lainey2016quantification}].
\textcolor{Corr}{In other studies, $3k_{\rm qp} / 2 k_{\rm p}$ is often notated
$k_2 / 2C$ \citep[e.g.][]{millholland_disk}}. In Eq.~\eqref{eq:wld}, $n \equiv
\sqrt{GM_\star/a_{\rm p}^3}$ is the planet's orbital mean motion, and we have
assumed $r_d\gg a_p$ and included only the leading-order (quadrupole)
interaction between the planet and disk. We define three relative inclination
angles via
\begin{equation}
  \uv{s} \cdot \uv{l}\equiv \cos\theta,\quad
  \uv{s} \cdot \uv{l}_{\rm d}\equiv \cos\theta_{\rm sd},\quad
  \uv{l} \cdot \uv{l}_{\rm d}\equiv \cos I.
\end{equation}
In our model, $I$ is a constant.
Following standard notation \citep[e.g.][]{colombo1966,peale1969,ward2004I}, we
have defined $\alpha \equiv \omega_{\rm sl}$ and $g \equiv -\omega_{\rm ld}\cos I$.

@@

When a trajectory experiences separatrix crossing, it transitions into nearby
zones probabilistically. This process is studied \textcolor{Corr}{in the
adiabatic limit} by \citet{henrard1982} and \citet{henrard1987}. Their results
may be summarized as follows: if zone $i$ is shrinking while adjacent zones $j,
k$ are expanding such that the sum of their areas is constant, the probabilities
of transition from zone $i$ to zones $j$ and $k$ are given by\footnote{
\textcolor{Corr}{Interestingly, the probabilistic nature of separatrix crossing
is not because the outcome depends on the phase $\phi$ of the system at
crossing. Instead, the transition the system takes is determined by the value of
$H$ at the start of a separatrix-crossing orbit \citep{henrard1982}. Since $H$
only changes by $\propto \epsilon$ over a circulation/libration cycle, $\sim
\epsilon$ changes to initial conditions affect the final outcome. As such, as
$\epsilon \to 0$, the outcome effectively becomes probabilistic.}}
\begin{subequations}\label{eq:henrard_hop}
    \begin{align}
        \Pr\p{i \to j} = -\frac{\pdil{\mathcal{A}_{\rm
            j}}{\eta}}{\pdil{\mathcal{A}_{\rm i}}{\eta}}, \\
        \Pr\p{i \to k} = -\frac{\pdil{\mathcal{A}_{\rm
            k}}{\eta}}{\pdil{\mathcal{A}_{\rm i}}{\eta}}.
    \end{align}
\end{subequations}
Note that $\Pr \p{i \to j} + \Pr\p{i \to k} = 1$.
Eqs.~\eqref{eq:henrard_hop} can be used in conjunction with
Eqs.~\eqref{se:area_ward} to understand for what initial conditions each track
can be observed and with what probabilities.

@@

As $\epsilon$ increases further (see Fig.~\ref{fig:3_ensemble_05_15}) but still
marginally satisfies the weak adiabaticity criterion [Eq.~\eqref{eq:ad_constr}],
the scatter in $\theta_{\rm f}$ continues to widen. The horizontal banded
structure of the final obliquities is a consequence of even stronger phase
sensitivity during separatrix crossing: trajectories cross the separatrix at
similar phases evolve to similar final obliquities that only depend weakly on
on $\theta_{\rm sd, i}$. \textcolor{Corr}{Finally, the bottom edge
of the data and the III $\to$ I track deviate very noticeably. This
non-adiabatic effect is the result of the separatrix evolving significantly
within the separatrix-crossing orbit, as the tracks computed in
Section~\ref{s:ad} assume $\eta$ is constant throughout the separatrix-crossing
orbit.}

@@

    \item As the disk dissipates more rapidly, the adiabatic condition
        [Eq.~\eqref{eq:ad_constr}] breaks down, we find that a broad range of
        final obliquities can be reached for a given $\theta_{\rm sd, i}$ (see
        Fig.~\ref{fig:nonad_3_ensemble}). We \textcolor{Corr}{understand this
        via the novel concept of partial adiabatic resonance advection and}
        provide an analytical expression of the bounds on $\theta_{\rm f}$ in
        Eq.~\eqref{eq:nonad_q_f_dist}.
\end{enumerate}

@@

\textcolor{Corr}{In practice, planets are not expected to form in isolation.
\citet{millholland_disk} point out that one effect of a companion planet is that
$g$ for the planet of interest never decays below the Laplace-Lagrange
secular nodal precession frequency. Therefore, $\eta$ has a minimum value at
late times. This does not affect the methodology of our analysis, but has some
small effects on our detailed results: (i) the adiabaticity criterion must be
modified slightly as $\rdil{\ln \eta}{t}$ is no longer constant but asymptotes
to zero as $\eta$ decreases, (ii) the planet may never undergo separatrix
crossing if their $\eta_{\star}$ (which depends on $\theta_{\rm sd, i}$) in the
absence of the companions is too small, and (iii) the planetary obliquities will
oscillate even when the disk has fully evaporated (as $\uv{l}$ is no longer
constant). An interesting consequence of (ii) is that an outer planet may
experience separatrix crossing before an inner planet does. The dynamics can be
even more complex if the two planets are in mean motion resonance
\citep[e.g.][]{millholland2019obliquity}.}

@@

We thank \textcolor{Corr}{the anonymous reviewer for their detailed and
insightful comments that improved this work and} Kassandra Anderson for
discussion and assistance in the early phase of this work. DL thanks the Dept.\
of Astronomy and the Miller Institute for Basic Science at UC Berkeley for
hospitality while part of this work was carried out. This work has been
supported in part by the NSF grant AST-17152 and NASA grant 80NSSC19K0444. YS is
supported by the NASA FINESST grant 19-ASTRO19-0041.% chktex 8

>>>>

In this paper, we consider a wide range of initial planetary obliquities in the
Millholland-Batygin dissipating disk scenario, and examine how the obliquity
evolves toward the ``final'' value as the exterior disk dissipates. We provide
an analytical framework for understanding the final planetary obliquity for
arbitrary initial spin-disk misalignment angles. We also consider various
dissipation timescales, and examine both ``adiabatic'' (slow disk dissipation)
and ``non-adiabatic'' evolution. We calibrate these analytical results with
numerical calculations. \textcolor{Corr}{ On the technical side, our paper
extends previous works \citep[such as][]{henrard1982, henrard1987,
millholland_disk} in several aspects. Two of our main results are: (i) a careful
accounting of the phase space area across sepratrix to analytically describe the
rich dynamics of adiabatic evolution, and (ii) using the concept of ``partial
adiabatic resonance advection'' to fully capture the dynamics in the
non-adiabatic limit.}

@@

The spin axis $\uv{s}$ of the planet tends to precess around its orbital
(angular momentum) axis $\uv{l}$, driven by the gravitational torque from the
host star acting on the planet's rotational bulge. On the other hand, $\uv{l}$
and the disk axis $\uv{l}_{\rm d}$ precess around each other due to
gravitational interactions. We assume $S \ll L \ll L_{\rm d}$, so $\uv{l}_{\rm
d}$ is nearly constant and $\uv{l}$ experiences negligible backreaction torque
from $\uv{s}$. The equations of motion for $\uv{s}$ and $\uv{l}$ in this limit
are \textcolor{Corr}{\citep{anderson2018teeter}}
\begin{align}
    \rd{\uv{s}}{t} &= \omega_{\rm sl} \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}}
        \equiv \alpha \p{\uv{s} \cdot \uv{l}}
            \p{\uv{s} \times \uv{l}},\label{eq:dsdt}\\
    \rd{\uv{l}}{t} &= \omega_{\rm ld}\p{\uv{l} \cdot \uv{l}_{\rm d}}
            \p{\uv{l} \times \uv{l}_{\rm d}}
        \equiv -g\p{\uv{l} \times \uv{l}_{\rm d}},
            \label{eq:dldt}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &\equiv \frac{3GJ_2 M_pR_p^2 M_\star}{2a_p^3 I_p\Omega_p}
        = \frac{3k_{\rm qp}}{2k_{\rm p}} \frac{M_\star}{m_{\rm p}}
            \p{\frac{R_{\rm p}}{a_{\rm p}}}^3 \Omega_{\rm p},\label{eq:wsl}\\
    \omega_{\rm ld} &\equiv \frac{3M_{\rm d}}{4M_\star}\p{\frac{a_{\rm
            p}}{r_{\rm d}}}^3 n .\label{eq:wld}
\end{align}
In Eq.~\eqref{eq:wsl}, $I_p = k_p M_pR_p^2$ (with $k_p$ a constant) is the
moment of inertia and $J_2=k_{\rm qp}\Omega_p^2 (R_p^3/GM_p)$ (with $k_{qp}$ a
constant) the rotation-induced (dimensionless) quadrupole of the planet [for a
body with uniform density, $k_p=0.4, k_{qp}=0.5$; for giant planets, $k_p\simeq
0.25$ and $k_{qp}\simeq 0.17$ \citep[e.g.][]{lainey2016quantification}].
\textcolor{Corr}{In other studies, $3k_{\rm qp} / 2 k_{\rm p}$ is often notated
as $k_2 / 2C$ \citep[e.g.][]{millholland_disk}}. In Eq.~\eqref{eq:wld}, $n
\equiv \sqrt{GM_\star/a_{\rm p}^3}$ is the planet's orbital mean motion, and we
have assumed $r_d\gg a_p$ and included only the leading-order (quadrupole)
interaction between the planet and disk. We define three relative inclination
angles via
\begin{equation}
  \uv{s} \cdot \uv{l}\equiv \cos\theta,\quad
  \uv{s} \cdot \uv{l}_{\rm d}\equiv \cos\theta_{\rm sd},\quad
  \uv{l} \cdot \uv{l}_{\rm d}\equiv \cos I.
\end{equation}
In our model, $I$ is a constant.
Following standard notation \citep[e.g.][]{colombo1966,peale1969,ward2004I}, we
have defined $\alpha \equiv \omega_{\rm sl}$ and $g \equiv -\omega_{\rm ld}\cos I$.

@@

When a trajectory experiences separatrix crossing, it transitions into nearby
zones probabilistically. This process is studied \textcolor{Corr}{in the
adiabatic limit} by \citet{henrard1982} and \citet{henrard1987}. Their results
may be summarized as follows: if zone $i$ is shrinking while adjacent zones $j,
k$ are expanding such that the sum of their areas is constant, the probabilities
of transition from zone $i$ to zones $j$ and $k$ are given by\footnote{
\textcolor{Corr}{Interestingly, the probabilistic nature of separatrix crossing
is not because the outcome depends on the phase $\phi$ of the system at
crossing. Instead, the transition the system takes is determined by the value of
$H$ at the start of a separatrix-crossing orbit \citep{henrard1982}. Since $H$
only changes by $\propto \epsilon$ over a circulation/libration cycle, an order
$\sim \epsilon$ change to the initial condition affects the final outcome. As
such, as $\epsilon \to 0$, the outcome effectively becomes probabilistic.}}
\begin{subequations}\label{eq:henrard_hop}
    \begin{align}
        \Pr\p{i \to j} = -\frac{\pdil{\mathcal{A}_{\rm
            j}}{\eta}}{\pdil{\mathcal{A}_{\rm i}}{\eta}}, \\
        \Pr\p{i \to k} = -\frac{\pdil{\mathcal{A}_{\rm
            k}}{\eta}}{\pdil{\mathcal{A}_{\rm i}}{\eta}}.
    \end{align}
\end{subequations}
Note that $\Pr \p{i \to j} + \Pr\p{i \to k} = 1$.
Eqs.~\eqref{eq:henrard_hop} can be used in conjunction with
Eqs.~\eqref{se:area_ward} to understand for what initial conditions each track
can be observed and with what probabilities.

@@

As $\epsilon$ increases further (see Fig.~\ref{fig:3_ensemble_05_15}) but still
marginally satisfies the weak adiabaticity criterion [Eq.~\eqref{eq:ad_constr}],
the scatter in $\theta_{\rm f}$ continues to widen. The horizontal banded
structure of the final obliquities is a consequence of even stronger phase
sensitivity during separatrix crossing: trajectories cross the separatrix at
similar phases evolve to similar final obliquities that only depend weakly on
on $\theta_{\rm sd, i}$. \textcolor{Corr}{Finally, in
Fig.~\ref{fig:3_ensemble_05_15}, the bottom edge of the data and the III $\to$ I
track deviate very noticeably. This non-adiabatic effect is the result of the
separatrix evolving significantly within the separatrix-crossing orbit, as the
tracks computed in Section~\ref{s:ad} assume that $\eta$ is constant throughout
the separatrix-crossing orbit.}

@@

    \item As the disk dissipates more rapidly, the adiabatic condition
        [Eq.~\eqref{eq:ad_constr}] breaks down, we find that a broad range of
        final obliquities can be reached for a given $\theta_{\rm sd, i}$ (see
        Fig.~\ref{fig:nonad_3_ensemble}). We \textcolor{Corr}{understand this
        result via the novel concept of partial adiabatic resonance advection
        and} provide an analytical expression of the bounds on $\theta_{\rm f}$
        in Eq.~\eqref{eq:nonad_q_f_dist}.
\end{enumerate}

@@

\textcolor{Corr}{Concerning the production of planetary obliquity with a
dissipating disk, when there are multiple planets in a system, the nodal
precession rate $g$ for the planet of interest never decays below the
Laplace-Lagrange rate driven by planet-planet secular interactions
\citep{millholland_disk}. Therefore, $\eta$ has a minimum value at late times.
This does not affect the methodology of our analysis, but can affect the
detailed results. For example, the adiabaticity criterion must be modified
slightly as $\rdil{\ln \eta}{t}$ is no longer constant but asymptotes to zero as
$\eta$ decreases; the planet may never undergo separatrix crossing if their
$\eta_{\star}$ (which depends on $\theta_{\rm sd, i}$) in the absence of the
companions is too small; the planetary obliquity will oscillate even when the
disk has fully evaporated (as $\uv{l}$ is no longer constant). The spin dynamics
can be even more complex if the two planets are in mean motion resonance
\citep[e.g.][]{millholland2019obliquity}.}

@@

We thank \textcolor{Corr}{the anonymous reviewer for detailed comments that
improved this work and} Kassandra Anderson for discussion and assistance in the
early phase of this work. DL thanks the Dept.\ of Astronomy and the Miller
Institute for Basic Science at UC Berkeley for hospitality while part of this
work was carried out. This work has been supported in part by the NSF grant
AST-17152 and NASA grant 80NSSC19K0444. YS is supported by the NASA FINESST
grant 19-ASTRO19-0041.% chktex 8

### 2020-08-18 writeups/Su_weak_tides.tex
done response

% label C_\pm on fig 2
% unify 3 & 4, choose consistent q/cos q
    \documentclass[
        fleqn,
        usenatbib,
        referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
        caption,
        subcaption,
    }
    \usepackage[T1]{fontenc}
    \captionsetup{compatibility=false}

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

\title[Analytical Exoplanet Obliquities]{Analytical Predictions of Explanet
Obliquities Generated by Planet-Disk Interactions}
\author[Y. Su and D. Lai]{
Yubo Su$^1$,
Dong Lai$^{1,2}$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
\\
$^2$ Tsung-Dao Lee Institute, 800 Dongchuan Road, Shanghai, 200240, China
}

\date{Accepted XXX\@. Received YYY\@; in original form ZZZ}

\pubyear{2020}

\begin{document}\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

@@

\begin{keywords}
planet--star interactions % chktex 8
\end{keywords}

@@

@@

Plots of CS locations as always in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{CS locations, TODO change axis labels etc.}\label{fig:cs_locs}
\end{figure}

@@

Plot of level curves of Hamiltonian as before, including $\mathcal{C}_{\pm}$
notation in \autoref{fig:1contours}.
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/0_eta/1contours.png}
    \caption{TODO label $\mathcal{C}_{\pm}$.}\label{fig:1contours}
\end{figure}

@@

The phase portrait for these EOM in $\p{s, \theta}$ space is shown in
Fig.~\ref{fig:quiver}.
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/2quiver0_2.png}
    \caption{Phase portrait.}\label{fig:quiver}
\end{figure}

@@

Generally, these EOM have equilibria at the points that are both a CS and
satisfy $\dot{s} = 0$ under weak tidal friction. These are stable as $\epsilon
\to 0$, see Appendix~\ref{app:cs_stab}. On Fig.~\ref{fig:quiver}, these are the
intersection of the locations of CS1 and CS2 with the line $\dot{s} = 0$. It is
clear that the locations of these equilibria depend on the value of $s_c$. Call
these generalized equilibria \emph{tidal Cassini Equilibria} (tCE), and number
them tCE1 and tCE2 depending on whether they are CS1 or CS2 states. Shown
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/6equils0_06.png}
    \caption{Location of tCE\@. Vertical line is spin rate below which tCE is
    destroyed.}\label{fig:6equils}
\end{figure}

@@

As a result of \autoref{ss:tidal_eqs}, the tCE are the only possible final
outcomes. We generate initial spin vectors $\hat{s}$ for isotropic initial
distribution $\p{\cos \theta, \phi}$, and marginalize over $\phi$ (which
generally cannot be fixed by any formation mechanism) in histograms. This gives
histograms in Figs.~\ref{fig:Hhists_0_06,fig:Hhists_0_20,fig:Hhists_0_70}
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5Hhists0_06_5.png}
    \caption{$s_c = 0.06$. Note that it is difficult to reach
    tCE2.}\label{fig:Hhists_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5Hhists0_20_5.png}
    \caption{$s_c = 0.2$. Note that tCE2 is both reached with substantial
    probability and has substantial obliquity.}\label{fig:Hhists_0_20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5Hhists0_60_5.png}
    \caption{$s_c = 0.6$. Note that tCE2 becomes attracting over tCE1, but will
    have obliquity $\approx I$ and is uninteresting.}\label{fig:Hhists_0_70}
\end{figure}

The governing mechanism for these probabilities is probabilistic separatrix
crossing, which is covered in Appendix~\ref{app:probs}. The calculation is
difficult analytically, but can be performed semi-analytically (describe
procedure here, calculate $\eta_{\star}$ numerically etc.). The agreement of
this curve with the histograms is examined for the same three $s_c$ values
in Figs.~\ref{fig:pc_fits_0_06,fig:pc_fits_0_20,fig:pc_fits_0_70}.
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5pc_fits0_06_5.png}
    \caption{$s_c = 0.06$ prediction vs histogram.}\label{fig:pc_fits_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5pc_fits0_20_5.png}
    \caption{$s_c = 0.20$ prediction vs histogram.}\label{fig:pc_fits_0_20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5pc_fits0_70_5.png}
    \caption{$s_c = 0.70$ prediction vs histogram.}\label{fig:pc_fits_0_70}
\end{figure}

@@

Assuming that $\hat{s}$ is drawn from an isotropic distribution, we may then
calculate the probabilities of going to either tCE as a function of $s_c$. This
is shown below for $I = 5^\circ$
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../initial/1_weaktide/5probs_5.png}
    \caption{Top: total probability of ending up in tCE2 (blue dots) and the
    prediction ignoring separatrix capture (red line). Bottom:
    obliquities of the tCEs.}\label{fig:probs}
\end{figure}

@@

\bibliographystyle{mnras}
\bibliography{Su_sep_cross}

\appendix
\section{Proof of Stability of Cassini States 1 \& 2}\label{app:cs_stab}

Insert here.

@@

\bsp
\label{lastpage} % chktex 24
\end{document}

>>>>

% label C_\pm on fig 2
% unify 3 & 4, choose consistent q/cos q
    \documentclass[twocolumn]{aastex63}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
    }
    \usepackage[T1]{fontenc}

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rdil}[2]{\mathrm{d}#1 / \mathrm{d}#2}
    \newcommand*{\pdil}[2]{\partial#1 / \partial#2}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\bsmb}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\uv}[1]{\hat{\bsmb{#1}}}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

    \newcommand\aastex{AAS\TeX}
    \newcommand\latex{La\TeX}

    \colorlet{Corr}{red}
    \received{XXXX}
    \revised{XXXX}
    \accepted{XXXX}
    \submitjournal{ApJ}

\shorttitle{Weak Tides and Cassini States}
\shortauthors{Y.\ Su and D.\ Lai}

\begin{document}

\title{Dynamics of Colombo's Top: Tidal Dissipation and Resonance Capture}

\correspondingauthor{Yubo Su}
\email{yubosu@astro.cornell.edu}

\author[0000-0001-8283-3425]{Yubo Su}% chktex 8
\affiliation{Cornell Center for Astrophysics and Planetary Science, Department
of Astronomy, Cornell University, Ithaca, NY 14853, USA}

\author[0000-0002-1934-6250]{Dong Lai}% chktex 8
\affiliation{Cornell Center for Astrophysics and Planetary Science, Department
of Astronomy, Cornell University, Ithaca, NY 14853, USA}

@@

\keywords{planet---star interactions}

@@

In this section, we first briefly lay out the spin dynamics of the planet,
introducing the Cassini State spin-orbit resonance \citep[for more details,
see][]{su2020}. We then introduce the equilibrium tide model of weak tidal
friction used in this work, though much of our analysis does not depend on the
detailed tidal model used.

\subsection{Spin Dynamics Without Tidal Friction}\label{ss:theory_spin}

@@

Plots of CS locations as always in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{CS locations, TODO change axis labels etc.}\label{fig:cs_locs}
\end{figure}

@@

Plot of level curves of Hamiltonian as before, including $\mathcal{C}_{\pm}$
notation in \autoref{fig:1contours}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/0_eta/1contours.png}
    \caption{TODO label $\mathcal{C}_{\pm}$.}\label{fig:1contours}
\end{figure}

@@

The phase portrait for these EOM in $\p{s, \theta}$ space is shown in
Fig.~\ref{fig:quiver}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/2quiver0_2.png}
    \caption{Phase portrait.}\label{fig:quiver}
\end{figure}
Note that the semimajor axis $a$ is not constant, but $\dot{a} / \dot{\theta}
\sim S / L \ll 1$, and is negligible for our problem (but not always,
Millholland tidal runaway cite).

@@

Generally, these EOM have equilibria at the points that are both a CS and
satisfy $\dot{s} = 0$ under weak tidal friction. These are stable when ignoring
tidal friction \citep{su2020}. On Fig.~\ref{fig:quiver}, these are the
intersection of the locations of CS1 and CS2 with the line $\dot{s} = 0$. It is
clear that the locations of these equilibria depend on the value of $s_c$. Call
these generalized equilibria \emph{tidal Cassini Equilibria} (tCE), and number
them tCE1 and tCE2 depending on whether they are CS1 or CS2 states. Shown
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/6equils0_06.png}
    \caption{Location of tCE\@. Vertical line is spin rate below which tCE is
    destroyed.}\label{fig:6equils}
\end{figure}

@@

As a result of \autoref{ss:tidal_eqs}, the tCE are the only possible final
outcomes. We generate initial spin vectors $\hat{s}$ for isotropic initial
distribution $\p{\cos \theta, \phi}$, and average over $\phi$ in histograms.
This gives histograms in
Figs.~\ref{fig:Hhists_0_06,fig:Hhists_0_20,fig:Hhists_0_70}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_06_5.png}
    \caption{$s_c = 0.06$. Note that it is difficult to reach
    tCE2.}\label{fig:Hhists_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_20_5.png}
    \caption{$s_c = 0.2$. Note that tCE2 is both reached with substantial
    probability and has substantial obliquity.}\label{fig:Hhists_0_20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_60_5.png}
    \caption{$s_c = 0.6$. Note that tCE2 becomes attracting over tCE1, but will
    have obliquity $\approx I$ and is uninteresting.}\label{fig:Hhists_0_70}
\end{figure}

Similarly to \citet{su2020}, these probabilities are the result of separatrix
crossings. However, the calculation differs from \citet{su2020}: since the
perturbation is dissipative in phase space (i.e.\ it does not conserve phase
space area), analysis following \citet{henrard1982} is not sufficient. A more
general theory of separatrix crossing indeed lets us semi-analytically calculate
the outcome probabilities. Appendix~\ref{app:probs} gives a more thorough
overview of separatrix crossing dynamics, but the key result is:
\begin{equation}
    a^2 + b^2 = c^2.
\end{equation}
We can apply this curve to calculate the likelihood of the tCEs when a
particular trajectory encounters the separatrix, and from this obtain a
semi-analytical prediction of the distribution of outcomes. The semi-analytical
nature of this calculation arises because $\eta_\star$ cannot be calculated
\emph{a priori}. The agreement of this curve with the histograms is examined for
the same three $s_c$ values in
Figs.~\ref{fig:pc_fits_0_06,fig:pc_fits_0_20,fig:pc_fits_0_70}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5pc_fits0_06_5.png}
    \caption{$s_c = 0.06$ prediction vs histogram.}\label{fig:pc_fits_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5pc_fits0_20_5.png}
    \caption{$s_c = 0.20$ prediction vs histogram.}\label{fig:pc_fits_0_20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5pc_fits0_70_5.png}
    \caption{$s_c = 0.70$ prediction vs histogram.}\label{fig:pc_fits_0_70}
\end{figure}

@@

Assuming that $\hat{s}$ is drawn from an isotropic distribution, we may then
calculate the probabilities of going to either tCE as a function of $s_c$. This
is shown below for $I = 5^\circ$
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5probs_5.png}
    \caption{Top: total probability of ending up in tCE2 (blue dots) and the
    prediction ignoring separatrix capture (red line). Bottom:
    obliquities of the tCEs.}\label{fig:probs}
\end{figure}

@@

\bibliography{Su_weak_tides}
\bibliographystyle{aasjournal}

\appendix

@@

\end{document}
